Convergent Evolution During Local Adaptation to Patchy Landscapes

### 2014-11-30 patchy-selection-paper.tex
updated with more sims

Here we make a more precise argument to back up expression \eqref{eqn:twocase_migrate} for the migration rate.
The argument made above in \nameref{ss:heuristics} applies to general migration mechanisms,
since it relies only on a decomposition of the migrant families upon hitting the new patch;
but it is also imprecise in a number of subtle ways
that are difficult to formalize.
The argument we make here therefore takes a somewhat different tack,
supposing that it suffices to model the spatial movement of a migrant family
by following only the motion of the ``trunk''
(i.e.\ the red line in figure \ref{fig:branching_decomp}),
and supposing this motion is Brownian, with variance $\sigma^2$ per generation.
(Recall $\sigma$ is the dispersal distance.)
We then use facts about Brownian motion and branching processes, 
to compute more precise versions of equations \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}.
\linelabel{rr:moved_argument}

@@

Intuitively, $h_A(x,y)$ counts each family once if it hits the patch,
while the occupation density $u(x,y)$ counts the total amount of time that the family spends in the patch.
Since the expected total time spent in the patch does not depend on the distance at which the family started
if we know that it survived to hit the patch,
it is intuitively reasonable that $u(x,y)$ and $h_A(x,y)$ are simply related.
Following this intuition,
and that $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
it turns out (see section~\nameref{ss:hitting_occupation}) that
\begin{align} 
    h_A(x,y) \approx \frac{ u(x,y) }{ \E[K] g(A) } ,
\end{align}
where $g(A)$ is the total expected occupation time of the patch given the family arrives.
Recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
Since the integrands in expressions \ref{eqn:occupation_integral} and \ref{eqn:migrate_integral} only differ by a factor
that (at least asymptotically) does not depend on the distance between $x$ and $y$,
we can obtain the migration rate by multiplying the equilibrium frequency by this factor.
The probability $f(A)$ that the family does not die out once it arrives at the new patch is
approximately $\E[1-(1-p_e)^K] \approx 2 s_p \E[K] / \xi^2$,
if each member arriving has chance $p_e$ of establishing, independently.
\plr{And if $p_e$ is small?}
% If we let $w$ be the width of the patch, then the remaining term is
% \begin{align}
%   s(w) \simeq \frac{1}{2s_m} \left( 1 - \left( \frac{\pi w \sqrt{2 s_m}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2s_m}/\sigma} \right).
% \end{align}
Combining with the fact that $g(A)\approx 1/(2s_m)$ (see section~\nameref{ss:hitting_occupation}, in particular equation \eqref{eqn:q_migrate_relation}),
we obtain that XXX FIX
\begin{align}
  \label{eqn:orig_migrate} 
  \migrate(y) &\approx q(y) \times \frac{ 4 \rho s_p  s_m }{ \xi^2 }  \\
  &\approx C \frac{ 4 \rho s_p s_m }{ \xi^2 } \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma) .
\end{align}
This does not depend on $A$, because although
$f(A)$ and $g(A)$ in principle depend on the patch size (and geometry),
the dependence is very weak.
For instance, the width of a circular patch only very weakly affects the chance 
of establishment of a new migrant that appears on its edge,
as long as the patch is large enough.
For more discussion, including an expression for the outflux of migrants $\gamma(x)$, see section~\nameref{ss:hitting_occupation}.

@@

If we let $q(S) = \int_S q(y) dy$ be the total equilibrium occupation of $S$,
then by equations~\eqref{eqn:occupation_integral} and \eqref{eqn:migrate_integral},
\begin{align}
    q(S)  &= \int_{\partial B(r_0)} \gamma(x) u(x,S) dx  \\
          &\approx g(A) \E[K] \int_{\partial B(r_0)} \gamma(x) h(x,S) dx \\
          &= \frac{ g(A) \E[K] }{ \rho f(A) } \migrate(y) . \label{eqn:q_migrate_relation}
\end{align}
The function $g(A)$ is the expected amount of time that a Brownian motion begun on the edge of a disk of area $A$
is expected to spend inside the disk before $\tau_\dagger$.
This is integral of the Green function for the Bessel process of the appropriate order, 
so using \citet{borodin2002handbook}, % (Appendix 1),
and letting $w$ be the width of the patch,
in $d=1$,
\begin{align} \label{eqn:gA_1D}
    g(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
    % &= \int_0^{2w} G_{s_m}(0,y) dy \\
  &= \int_0^{2w/\sigma} \frac{ e^{- y \sqrt{2s_m}/\sigma }}{\sqrt{2s_m}} dy \\
  &= (1-e^{-2w\sqrt{2s_m/\sigma}})/(2s_m)
\end{align}
and in $d=2$, by \citet{gradshteyn2007table} 5.56.2,
\begin{align}
    g(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
    % &= \int_0^{2w} G_{s_m}(0,y) dy \\
  &= \int_0^{2w/\sigma} 2 y K_0(y \sqrt{2s_m}) dy \\
  &= \frac{1}{s_m}\int_0^{2w\sqrt{2s_m}/\sigma} y K_0(y) dy \\
  &= \frac{1}{2s_m}\left(1- \frac{2w\sqrt{2s_m}}{\sigma} K_1(2w\sqrt{2s_m}/\sigma) \right)  .
\end{align}
In either case, $g(A) \approx 1/(2s_m)$, which is the approximation we use in the main text.

>>>>

Here we make a more precise argument to back up expression \eqref{eqn:migrate} for the migration rate.
The argument made above in \nameref{ss:heuristics} applies to general migration mechanisms,
since it relies only on a decomposition of the migrant families upon hitting the new patch;
but it is also imprecise in a number of subtle ways
that are difficult to formalize.
Here we take a somewhat different tack,
supposing that it suffices to model the spatial movement of a migrant family
by following only the motion of the ``trunk''
(i.e.\ the red line in figure \ref{fig:branching_decomp}),
and supposing this motion is Brownian, with variance $\sigma^2$ per generation.
(Recall $\sigma$ is the dispersal distance.)
We then use facts about Brownian motion and branching processes, 
to compute more precise versions of equations \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}.
\linelabel{rr:moved_argument}

@@

Intuitively, $h_A(x,y)$ counts each family once if it hits the patch,
while the occupation density $u(x,y)$ counts the total amount of time that the family spends in the patch.
Since the expected total time spent in the patch does not depend on the distance at which the family started
if we know that it survived to hit the patch,
it is intuitively reasonable that $u(x,y)$ and $h_A(x,y)$ are simply related.
Following this intuition,
and that $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
it turns out (see section~\nameref{ss:hitting_occupation}) that
\begin{align} 
    h_A(x,y) \approx \frac{ u(x,y) }{ \E[K] g(A) } ,
\end{align}
where $g(A)$ is the total expected occupation time of the patch given the family arrives.
Recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
Since the integrands in expressions \ref{eqn:occupation_integral} and \ref{eqn:migrate_integral} only differ by a factor
that (at least asymptotically) does not depend on the distance between $x$ and $y$,
we can obtain the migration rate by multiplying the equilibrium frequency by this factor.

The result will not depend on $A$, 
because although $f(A)$ and $g(A)$ in principle depend on the patch size (and geometry),
the dependence is very weak.
For instance, the width of a circular patch only very weakly affects the chance 
of establishment of a new migrant that appears on its edge,
as long as the patch is large enough.
For more discussion, including an expression for the outflux of migrants $\gamma(x)$, see section~\nameref{ss:hitting_occupation}.

@@

If we let $q(S) = \int_S q(y) dy$ be the total equilibrium occupation of $S$,
then by equations~\eqref{eqn:occupation_integral} and \eqref{eqn:migrate_integral},
\begin{align}
    q(S)  &= \int_{\partial B(r_0)} \gamma(x) u(x,S) dx  \\
          &\approx g(A) \E[K] \int_{\partial B(r_0)} \gamma(x) h(x,S) dx \\
          &= \frac{ g(A) \E[K] }{ \rho f(A) } \migrate(y) . \label{eqn:q_migrate_relation}
\end{align}
Once we show that $g(A) \approx 1/(2 s_m)$, 
we will have arrived at the result, equation \eqref{eqn:ratio_K}.

The function $g(A)$ is the expected amount of time that a Brownian motion begun on the edge of a disk of area $A$
is expected to spend inside the disk before $\tau_\dagger$.
This is integral of the Green function for the Bessel process of the appropriate order, 
so using \citet{borodin2002handbook}, % (Appendix 1),
and letting $w$ be the width of the patch,
in $d=1$,
\begin{align} \label{eqn:gA_1D}
    g(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
    % &= \int_0^{2w} G_{s_m}(0,y) dy \\
  &= \int_0^{2w/\sigma} \frac{ e^{- y \sqrt{2s_m}/\sigma }}{\sqrt{2s_m}} dy \\
  &= (1-e^{-2w\sqrt{2s_m/\sigma}})/(2s_m)
\end{align}
and in $d=2$, by \citet{gradshteyn2007table} 5.56.2,
\begin{align}
    g(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
    % &= \int_0^{2w} G_{s_m}(0,y) dy \\
  &= \int_0^{2w/\sigma} 2 y K_0(y \sqrt{2s_m}) dy \\
  &= \frac{1}{s_m}\int_0^{2w\sqrt{2s_m}/\sigma} y K_0(y) dy \\
  &= \frac{1}{2s_m}\left(1- \frac{2w\sqrt{2s_m}}{\sigma} K_1(2w\sqrt{2s_m}/\sigma) \right)  .
\end{align}
In either case, $g(A) \approx 1/(2s_m)$, which is the approximation we use in the main text.
